LME-MPM applied to quasi-brittle fracture.

### 2020-05-18 2020_EFM_MPM_Eigensoftening.tex
Cambios 18-5-2020

@@

\newcommand{\vec}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}

@@

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the material point method (MPM)
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the optimal transportation meshfree (OTM)
  approximation scheme, is extended to the MPM with the aim of present
  a suitable alternative to the existing fracture algorithms developed
  for the MPM. The good fitting in the predictions made by the
  eigensoftening algorithm against both analytical and experimental
  results proofs the well performance of the method under challenging loads.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
Presence of cracks are a violation of the continuity requirement of
the finite element method (FEM). To overcome it, numerous of numerical
artifacts has been proposed with the aim of reproducing  such a
complex behaviour. This techniques vary from employing cohesive
approaches \cite{Barenblatt,Hilleborg_1976}, by adaptively inserting
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
techniques. Some examples of it are the material
point method (MPM) \cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the
element-free Galerkin method (EFGM) \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the smoothed particle hydrodynamics (SPH) \cite{Wang_2020,Wang_2019},
the optimal transportation meshfree method (OTM)
\cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or peridynamics
\cite{HA_2011,RABCZUK_2017} among others.

Fracture can be described in two ways. One is to remove the
restriction of the single-valued velocity field close to the crack by
using two or more sets of nodes \cite{Nairn_2006}. This method is
based in to assign different labels to distinguish if the material
points and nodes are in the same side of the crack or not. Under this
approach the crack surface is described with line segments in 2D and
triangle patches in 3D cases. This method was named as ``CRAcks with
Material Points (CRAMP)'' \cite{Nairn_2003}. In this method, the
criteria for crack propagation is based on such parameters as energy
release rate analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the
stress intensity factor or the J-integral discussed by Guo \& Nairn 
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter
its propagation and branching. Consequently, the position of the crack
does not need to be explicitly stated. This represents a significant
advantage over the ``CRAMP''. Under this approach, the prediction the
failure evolution employing a decohesion model has been discussed by
Chen {\it et al.}\cite{Zhenmao_2005} or Schreyer {\it et
  al.}\cite{Schreyer_2002}. And has been successfully employed to
simulate the fracture of brittle materials by Chen {\it et al.}
\cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}. Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
they approximate the crack set by means of eigen-deformations, wich
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle facture
was developed by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique the erosion of
the material point means that each material point can be either intact
or be completely failed or eroded and has no load bearing
capacity. This method has been successfully applied to simulate high
complex phenomena such dynamic fragmentation of metals \cite{Li_2015},
nonetheless in quasi-brittle materials, simulations performed with
this approach exhibit an overestimation of tensile stress and the
strain peaks. To overcome this limitation, Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} introduced the concept of
eigensoftening. Engineered for quasi-brittle materials, this approach
is devoted to capture the gradual rather than abrupt dissipation of
the fracture energy. It is inspired in the concept of the crack band
model \cite{Bazant83}, since energy dissipation is thorough the
softened (or failed) volume. Due to the similarities between the OTM, 
where this procedure was developed, and the MPM, we propose

The eigensoftening algorithm together with a explicit predictor
corrector and the local \textit{max-ent} approximation technique
offers a new perspective in the MPM fracture simulation.

@@

In the current work we attempt to model the fracture propagation
employing the MPM and introducing both eigenerosion and eigensoftening
\cite{Navas_2017_ES,Navas_2018_ES} and the concept of the
eigenfracture under this framework.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}
The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the MPM
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the
MPM will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}
The MPM~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the MPM possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the MPM,
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the MPM. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the MPM algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

(LME)$_\beta$ is,

@@

The regularization parameter $\beta$ of LME shape functions may be

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of MPM formulation, fracture can be modelled by
failing particles according to a suitable criterion. When material
points are failed, they are assumed to have a null stress
tensor. Navas {\it et al.} (2017)\cite{Navas_2017_ES} developed a
eigensoftening algorithm as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. A
comparison between both in \cite{Navas_2017_ES} shows that
the eigenerosion algorithm significantly overestimates the tensile
stress and the strain peaks, while it captures the forces and crack
patterns accurately. On the other hand eigensoftening algorithm agree
very well with experimental results in all the aspects. Furthermore,
this algorithm has also proof its accuracy for complex fracture
patters such the present in fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

\subsection{Comparison with analytical solution}
\label{sec:3.1}

\subsection{Brazilian test}
\label{sec:3.2}

\begin{figure}
  \centering
  \includegraphics[width=0.3\textwidth]{Figures/Brazilian}
  \caption{Geometry and boundary condition of the Brazilian test.}
  \label{fig:geometry-brazilian-test}
\end{figure}

@@

\section*{Acknowledgements}
The first author acknowledges the fellowship Agustín de Betancourt 262390106114.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init macros math

@@

\newcommand{\vect}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\tens}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Matrix}[1]{
  \ensuremath{\mathbf{{#1}}}
}
\newcommand{\Vector}[1]{
  \ensuremath{\mathbf{{#1}}}
}
% Divergence
\newcommand{\Div}[1]{
  \ensuremath{div({#1})}
}
% Gradient
\newcommand\Grad[1]{grad({#1})}
\newcommand\GradS[1]{grad^s({#1})}
\newcommand\GradT[1]{grad^T({#1})}
% Partial derivative
\newcommand{\Deriv}[3][]{
  \ensuremath{\frac{\partial^{#1}{#2}}{ \partial {#3}^{#1} }}
}
% Integral
\newcommand{\Integral}[2]{
  \IfStrEqCase{#1}{
    {2}{\ensuremath{\int_{\varGamma_d}{#2}\ d\varGamma}}
    {3}{\ensuremath{\int_{\varOmega}{#2}\ d\varOmega}}
  }
}
% End macros math
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Init glossaries
\usepackage[acronym]{glossaries}

\makeglossaries

\newglossaryentry{domain}
{
  name=$\varOmega$,
  description={Continuum domain}
}

\newglossaryentry{contour}
{
  name=$\partial\varOmega$,
  description={Boundary of the continuum domain $\varOmega$. Also
    defined as $\Gamma$}
}

\newglossaryentry{dirichlet-boundary}
{
  name=$ \Gamma_d$,
  description={Essential or Dirichlet boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{neumann-boundary}
{
  name=$ \Gamma_n$,
  description={Natural or Nemann boundary conditions over $\partial\varOmega$}
}

\newglossaryentry{rho}
{
  name=$\rho$,
  description={Describes the scalar density field}
}

\newglossaryentry{a}
{
  name=$\vect{a}$,
  description={First order tensor which describes the acceleration field}
}

\newglossaryentry{v}
{
  name=$\vect{v}$,
  description={First order tensor which describes the velocity field}
}

\newglossaryentry{u}
{
  name=$\vect{u}$,
  description={First order tensor which describes the displacement field}
}

\newglossaryentry{x}
{
  name=$\vect{x}$,
  description={First order tensor which describes the global coordinates field}
}

\newglossaryentry{xi}
{
  name=$\vect{\xi}$,
  description={First order tensor which describes the local coordinates field}
}

\newglossaryentry{stress}
{
  name=$\tens{\sigma}$,
  description={Second order tensor which means the Cauchy stress tensor}
}

\newglossaryentry{strain}
{
  name=$\tens{\varepsilon}$,
  description={Second order tensor which means the Cauchy strain tensor}
}

\newglossaryentry{Constitutive}
{
  name=$\tens{D}$,
  description={Four order tensor which means the constitutive response
    of the material}
}

\newglossaryentry{LME-beta}
{
  name=$\beta$,
  description={Regularization or thermalization parameter of the
    LME$_{\beta}$ Pareto set}
}

\newacronym{mpm}{MPM}{Material Point Method}
\newacronym{otm}{OTM}{Optimal Transportation Meshfree}
\newacronym{sph}{SPH}{Smoothed Particle Hydrodynamics}
\newacronym{fem}{FEM}{Finite Element Method}
\newacronym{efgm}{EFGM}{Element-Free Galerkin Method}
\newacronym{lme}{LME}{Local Maximum-Entropy}
\newacronym{flip}{FLIP}{Fluid Implicit Particle}
\newacronym{gimp}{GIMP}{Generalized Interpolation Material Point}
\newacronym{igimp}{iGIMP}{Implicit GIMP}
\newacronym{ugimp}{uGIMP}{Uniform GIMP}
\newacronym{ddmp}{DDMP}{Dual Domain Material Point}
\newacronym{ctls}{CTLS}{Conservative Taylor Least Squares}
\newacronym{npc}{NPC}{Newmark Predictor-Corrector}
\newacronym{fe}{FE}{Forward Euler}

% End glossaries
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{abstract}
  The objective of this work is to introduce an alternative
  technique to address the fracture process of brittle and
  quasi-brittle materials under the \acrfull{mpm} 
  framework. With this purpose the eigensoftening algorithm, developed
  originally for the \acrfull{otm} approximation scheme, is extended
  to the \acrshort{mpm} with the aim of present a suitable alternative to the
  existing fracture algorithms developed for the \acrshort{mpm}. The good fitting
  in the predictions made by the eigensoftening algorithm against both
  analytical and experimental results proofs the well performance of
  the method under challenging loads.
\end{abstract}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{sec:1}
Presence of cracks are a violation of the continuity requirement in
\acrfull{fem}. To overcome it, numerous of numerical
artifacts has been proposed with the aim of reproducing  such a
complex behaviour. This techniques vary from employing cohesive
approaches \cite{Barenblatt,Hilleborg_1976}, by adaptively inserting
cohesive elements \cite{Ortiz_1999,Pandolfi_2002,Ruiz_2000} at solid
elements boundaries, or handling arbitrary cracks paths by level set
representation of the fracture surface \cite{Belytschko_03}.

The simulation of fracture propagation in a more accurate and
effective way can be considered as one of the original drivers for
developing novel spatial discretization methods such as meshfree
techniques. Some examples of it are the \acrfull{mpm}
\cite{Schreyer_2002,Chen_2002,Nairn_2003,Zhenmao_2005}, the \acrfull{efgm}  \cite{BELYTSCHKO_1995,BELYTSCHKO_2000,Zhuang_2012,Muthu_2013},
the \acrfull{sph} \cite{Wang_2020,Wang_2019},
the \acrfull{otm} \cite{Li2010,Li_2012,Pandolfi_2013,Li_2015} or Peridynamics
\cite{HA_2011,RABCZUK_2017} among others.

Fracture can be described in two ways. One is to remove the
restriction of the single-valued velocity field close to the crack by
using two or more sets of nodes \cite{Nairn_2006}. This method is
based in to assign different labels to distinguish if the material
points and nodes are in the same side of the crack or not. Under this
approach the crack surface is described with line segments in 2D and
triangle patches in 3D cases. This method was named as ``CRAcks with
Material Points (CRAMP)'' \cite{Nairn_2003}. In this method, the
criteria for crack propagation is based on such parameters as energy
release rate analyzed by Tan \& Nairn (2002)\cite{Nairn_2002}, and the
stress intensity factor or the J-integral discussed by Guo \& Nairn 
(2004)\cite{Nairn_2004}. The other approach is to use failed material
points to describe the crack evolution. In this method, the formation
of failed points describes the nucleation of cracks, and thereafter
its propagation and branching. Consequently, the position of the crack
does not need to be explicitly stated. This represents a significant
advantage over the ``CRAMP''. Under this approach, the prediction the
failure evolution employing a decohesion model has been discussed by
Chen {\it et al.}\cite{Zhenmao_2005} or Schreyer {\it et
  al.}\cite{Schreyer_2002}. And has been successfully employed to
simulate the fracture of brittle materials by Chen {\it et al.}
\cite{Chen_2002,Chen_2003} and Sulsky \& Schreyer
\cite{Sulsky_2004}. Similar to this approach, Schmidt {\it et al.}
\cite{Schmidt_2009} introduced the concept of eigenfracture, where
they approximate the crack set by means of eigen-deformations, wich
enable the material to develop displacement jumps at no cost of local
elastic energy. Later, the eigenerosion approach to brittle facture
was developed by Pandolfi {\it et al.}
\cite{Pandolfi_2012,Pandolfi_2013}. In this technique the erosion of
the material point means that each material point can be either intact
or be completely failed or eroded and has no load bearing
capacity. This method has been successfully applied to simulate high
complex phenomena such dynamic fragmentation of metals \cite{Li_2015},
nonetheless in quasi-brittle materials, simulations performed with
this approach exhibit an overestimation of tensile stress and the
strain peaks. To overcome this limitation, Navas {\it et al.}
\cite{Navas_2017_ES,Navas2018a} introduced the concept of
eigensoftening. Engineered for quasi-brittle materials, this approach
is devoted to capture the gradual rather than abrupt dissipation of
the fracture energy. It is inspired in the concept of the crack band
model \cite{Bazant83}, since energy dissipation is thorough the
softened (or failed) volume. Due to the similarities between the \acrshort{otm}, 
where this procedure was developed, and the \acrshort{mpm}, we propose

The eigensoftening algorithm together with a explicit predictor
corrector and the local \textit{max-ent} approximation technique
offers a new perspective in the \acrshort{mpm} fracture simulation.

@@

In the current work we attempt to model the fracture propagation
employing the \acrshort{mpm} and introducing both eigenerosion and eigensoftening
\cite{Navas_2017_ES,Navas_2018_ES} and the concept of the
eigenfracture under this framework.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The meshfree methodology}
\label{sec:2}
The aim of this section is to provide an overview of the special
techniques employed to face the fracture problem under the \acrshort{mpm}
framework. In consequence it is structured as follows: first in
\ref{sec:2.1} the explicit-predictor algorithm for the \acrshort{mpm} will be exposed, next the local \textit{max-ent} approximants are introduced in
\ref{sec:2.2} as an accurate alternative technique to interpolate data
between particles and nodes, and finally the fracture algorithms based
in the eigendeformations are presented in \ref{sec:2.3}.

\subsection{The Material Point Method}
\label{sec:2.1}
The \acrshort{mpm}~\cite{Sulsky1994} is a meshfree Lagrangian-Eulerian method
where particles carries on all the physical information and a set of
background nodes is employed to compute the equilibrium
equation. Since the \acrshort{mpm} possesses the advantages of both Lagrangian
and Eulerian descriptions, no element distortion exists in the \acrshort{mpm},
therefore it is an appropriate and efficient method in solving
problems involving extremely large deformation and moving
discontinuities such fracture evolution. For the spatial
discretization, two sets of points are introduced in 
the \acrshort{mpm}. First, the nodes. This points are considered fixed in the
space and are in charge of computing all the kinematic fields such
forces $f_I$, accelerations $a_I$ and velocities $v_I$. And second the
material points or particles. They are in charge of the discretization
of the continuum, and store the local state ($\sigma_p,
\varepsilon_p$). Without loosing generality, the \acrshort{mpm} algorithm
can be described with three main steps: (i) a variational recovery
process, where particle data is projected to the grid nodes, (ii) an
Eulerian step, where balance of momentum equation is expressed as a
nodal equilibrium equation thorough a FEM-like procedure, and finally
(iii) a Lagrangian advection of the particles. In the present research
a explicit predictor-corrector time integration
scheme is adopted. The purpose of this choice is motivated due its proved
robustness and stability in numerical calculations. In the first stage
(i), the nodal velocity predictor is computed following
\eqref{eq:Predictor-velocity}, 
\begin{equation}
  \label{eq:Predictor-velocity}
  \vec{\tilde{v}}_I^{k+1} = \frac{ N_{Ip}^{k} m_p (\vec{v}_p^k + (1 - \gamma)\ \Delta t\ \vec{a}_p^k)}{m_I}
\end{equation}
This way of computing the nodal predictor is both numerically stable
and minimize the computational effort. Once nodal velocity are
obtained, the essential boundary conditions are imposed. And in the
following, a Eulerian phase (ii) is computed in the set of nodes in a
FEM-like way, where nodal forces $\vec{f}_{I}^{k+1}$ are computed thorough the
equilibrium equation. Next the nodal velocities are corrected in a
\textit{corrector} stage,
\begin{equation}
  \label{eq:Corrector-velocity}
  \vec{v}_{I}^{k+1} = \vec{v}_{I}^{pred} + \gamma\ \Delta t\ \frac{\vec{f}_{I}^{k+1}}{\tens{m}_I^{k+1}}
\end{equation}
Finally updated the particles are advected in the Lagrangian stage (iii) using nodal values as,
\begin{align}
  \label{eq:Update-lagrangian-pce}
        &\vec{a}_p^{k+1} = \frac{N_{Ip}^k\vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{v}_p^{k+1} = \vec{v}_p^n + \Delta t\
        \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k}\\
      &\vec{x}_p^{k+1} = \vec{x}_p^n + \Delta t\
         N_{Ip}^k\ \vec{v}_{I}^{k} +
        \frac{1}{2}\Delta t^2\ \frac{N_{Ip}^k\
        \vec{f}_{I}^{k}}{\tens{m}_I^k} 
\end{align}
The complete pseudo-algorithm it is summarized in \ref{sec:expl-pred-corr}.

@@

 \acrshort{lme}$_\beta$ is,

@@

The regularization parameter $\beta$ of \acrshort{lme} shape functions may be

@@

\subsection{Fracture modelling approach}
\label{sec:2.3}
Within the context of \acrshort{mpm} formulation, fracture can be modelled by
failing particles according to a suitable criterion. When material
points are failed, they are assumed to have a null stress
tensor. Navas {\it et al.} (2017)\cite{Navas_2017_ES} developed a
eigensoftening algorithm as an extension for quasi-brittle materials
of the eigenerosion proposed by Pandolfi \& Ortiz
(2012)\cite{Pandolfi_2012} for fracture of brittle materials. A
comparison between both in \cite{Navas_2017_ES} shows that
the eigenerosion algorithm significantly overestimates the tensile
stress and the strain peaks, while it captures the forces and crack
patterns accurately. On the other hand eigensoftening algorithm agree
very well with experimental results in all the aspects. Furthermore,
this algorithm has also proof its accuracy for complex fracture
patters such the present in fiber reinforces concrete (FRC),
\cite{Navas_2018_ES}.\\

@@

\subsection{Edge-cracked square panel in mode I}
\label{sec:edge-cracked-square-I}

This problem is devoted to compare the develop of both linear shape
functions and the \acrshort{lme}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figures/Mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[width=0.8\textwidth]{Figures/Reactions_displacements_square_panel_mode_I}
  \caption{Geometry and boundary condition of the drop-weight impact test.}
  \label{fig:geometry-cracked-panel-mode-I}
\end{figure}

@@

\section*{Acknowledgements}
The financial support to develop this research from the Ministerio de
Ciencia e Innovaci\'on, under Grant No. BIA-2016-76253 is greatly
appreciated. The first and the second authors also acknowledge the
fellowship Fundaci\'on Agust\'in de Betancourt and Juan de la Cierva
(FJCI-2017–31544) respectively.
